---
title: "Shiny Apps"
date: today
date-format: long
footer: "[DataViz 2102 Website](https://dataviz-gwu.rocks)"
logo: images/dataviz-logo.png
format:
  revealjs:
    theme: [simple, custom.scss]
    transition: fade
    slide-number: true
    multiplex: true
    chalkboard: true
execute:
  echo: true
  message: false
  warning: false
  freeze: auto
---

# Exploring Shiny 

## What is Shiny?

::: {.incremental}
- Shiny is an R package developed by [Posit](www.posit.co) that makes it easy to build interactive web apps straight from R
- No HTML, CSS, or JavaScript required
- Shiny apps can be hosted on the web or run locally
- Makes it easy to explore your data and share your analysis
:::

## What You Can Do with Shiny {.smaller}

::: {.incremental}
- **Data Exploration**: Utilize interactive widgets to filter and sort data dynamically
- **Basic Visualizations**: Create real-time, interactive graphs and charts from datasets
- **Advanced Analytics**: Integrate R's robust statistical and machine learning capabilities for in-depth analysis
- **Interactive Reports**: Build dynamic reports that update based on user input or live data
- **Production-Grade Dashboards**: Develop complex, multi-page dashboards with user authentication and data streaming
:::

## Flavors of Shiny {.smaller}

:::{.incremental}
- [R Shiny](https://shiny.posit.co/r/getstarted/shiny-basics/lesson1/index.html): The original framework, allowing R users to build interactive web applications directly within R
- [Shiny for Python](https://shiny.posit.co/py/): Extends Shiny's capabilities to Python, enabling Python users to create Shiny applications using Python code
- [Shiny Live](https://shiny.posit.co/py/docs/jupyter-widgets.html#shinylive): Enables you to run apps off the user's browser rather than having to deploy them to a server (only available in Python right now)
- [Shiny Express](https://shiny.posit.co/blog/posts/shiny-express/): A simplified version of Shiny for Python aimed at rapid development and deployment of applications with less coding and configuration required
- [Quarto Dashboards](https://quarto.org/docs/dashboards/): Not a Shiny app per se, but can connected to a Shiny server for increased interactivity
:::

## User Demos

- Have a look at the [Shiny gallery](https://shiny.posit.co/r/gallery/#user-showcase)
- Pick an app that you like?
- One thing you would like to learn from that app?
- One thing that you would improve
- Discuss with a neighbor

```{r}
#| echo: false
#| label: timer1

library(countdown)
countdown(minutes = 5, 
          id = "timer1", 
          bottom = "10%",
          right = "10%",
          color_border = "#fff",
          color_text = "#fff",
          color_running_background = "#42affa",
          color_running_text = "black",
          color_finished_background = "#E5D19D",
          color_finished_text = "#00264A")
```

## Feature Demos

- Now let's look at [feature demos](https://shiny.posit.co/r/gallery/#feature-demos)
- What widgets, layouts, or features are you interested in?
- How would you use them in your own apps?

```{r}
#| label: timer2
#| echo: false

countdown(minutes = 5, 
          id = "timer2", 
          bottom = "10%",
          right = "10%")
```

## Professional Apps {.smaller}

- Here is a selection of modern, professional Shiny Apps
  - [PESKAS](https://timor.peskas.org/) fisheries monitoring in Timor-Leste
  - [WaCSE](https://wsda.shinyapps.io/WaCSE/) for tracking reductions in GHG emissions as part of Washington State's [Sustainable Farms and Fields](https://www.scc.wa.gov/programs/sustainable-farms-fields) program
  - [JHU Lyme Disease Tracker](https://www.hopkinslymetracker.org/)
  - [movie-vue-r](https://friss.shinyapps.io/shiny-vue-r/#/) a Vue.js and Shiny app for exploring movie + COVID information
- These are mainly for inspiration, but try to find the code on GitHub and see how much you can understand

```{r}
#| label: timer3
#| echo: false

countdown(minutes = 5, 
          id = "timer3", 
          bottom = "10%",
          right = "10%")
```

## Getting Started with Shiny

- Open RStudio
- Start a folder for the workshop 
- Install the `shiny` package (if not installed already)
- Create a new Shiny app
  - (File > New File > Shiny Web App)
- Run the Geyser App
- Try opening it in a browser

```{r}
#| label: timer4
#| echo: false

countdown(minutes = 5, 
          id = "timer4", 
          bottom = "10%",
          right = "10%")
```

## Challenge

- Look at the `faithful` dataset in R (hint: `?faithful`)
- Create a Shiny app that displays a histogram of the `eruptions` column
- Change the labels accordingly
- What kind of distribution do you see?

```{r}
#| label: timer5
#| echo: false

countdown(minutes = 10, 
          id = "timer5", 
          bottom = "10%",
          right = "10%")
```

# Elements of a Shiny App

## Shiny App Basics

Two parts of a Shiny app:

- **UI**: User Interface
- **Server**: Server logic

They come together in a single script: `app.R`

::: aside
**Note:** You can compose your app with the ui and server in separate scripts, but we will stick to the single `app.R` script for now.
:::

## The app.R Script

```{r} 
#| label: app.r
#| code-line-numbers: "|1|3-5|7-8|10"
#| eval: false

library(shiny)

ui <- fluidPage(
  "Hello, Shiny!"
)

server <- function(input, output, session) {
}

shinyApp(ui, server)
```

# The User Interface (`ui`)

## The UI

```{r} 
#| label: ui
#| code-line-numbers: "3-5"
#| eval: false

library(shiny)

ui <- fluidPage(
  "Hello, Shiny!"
)

server <- function(input, output, session) {
}

shinyApp(ui, server)
```

`fluidPage()` creates a Bootstrap page

## The UI

- [Bootstrap](https://getbootstrap.com/docs/5.0/getting-started/introduction/) is a free front-end framework used in web development  
- It is built on **HTML**, **CSS** and **JavaScript**
- There are many version of Bootstrap, the latest being version 5, but R Shiny uses version 3 by default

::: aside
With the `bslib` package you can use Bootstrap 4 or 5 in your R Shiny apps.
:::

## The UI

```{r}
#| echo: false

library(shiny)
```

Under the hood, every UI function is translated to HTML.

```{r}
as.character(titlePanel("Hello, Shiny!"))
```

<br>

This HTML can have styles attached to it, either by the `style` attribute, or by a CSS class.

```{r}
as.character(sidebarPanel())
```

## The UI

To let users interact with your app, you can add [input controls](https://shiny.posit.co/r/getstarted/shiny-basics/lesson3/). A basic input control has:

- an `inputId`
- a `label`
- a `value`

## The UI

For example:

```{.r}
numericInput(inputId = "number", 
             label = "Enter a number", 
             value = 0)
```
<br>
```{.r}
actionButton(inputId = "button", 
             label = "Click me")
```

## Updating the UI

::: {.incremental}
- When the app runs, every user gets served the same HTML from the UI
- When the user interacts with the UI, we want that HTML to **react** and update based on user input
- That is possible because these outputs are **reactive outputs**
- The server logic (which we will discuss in detail soon) uses **reactive programming** that updates related inputs in the UI
:::

## Updating the UI

Shiny has several functions that turn R objects into reactive outputs for your `ui`: the [Output](https://shiny.posit.co/r/getstarted/shiny-basics/lesson4/) family of functions.

Each function creates a specific type of output, for example:

| UI Function | Output type |
|----------|-------------|
| textOutput() | text |
| tableOutput() | table |
| plotOutput() | plot |
| uiOutput() | raw HTML |
: {.table .table-striped .table-hover}

## Updating the UI

Every output element needs a single argument: `outputId`. This is a simple string that needs to be **unique**.

<br>

```{.r}
textOutput(outputId = "text")
```

<br>

This `textOutput` tells Shiny what to display. It's a **placeholder** for what is goign to be produced by the server logic.

# The Server Logic

## The Server

::: {.incremental}
- The server function builds an object named **output** and this object will contain all the code needed to display the R objects in the UI
- This output object is **list-like**
- Each R object that you want to display has its own entry in this list
- Therefore the name of every output in your UI needs to match a definition in the server, e.g. `output$text`
:::

## The Server

Each entry in the output list should contain a **Render** function. You must use **Render** functions that match the **Output** functions:

| UI Function | Output type | Render function |
|----------|-------------|-----------------|
| textOutput() | text | renderText() |
| tableOutput() | table | renderTable() |
| plotOutput() | plot | renderPlot() |
| uiOutput() | raw HTML | renderUI() |
: {.table .table-striped .table-hover}

## The Server

```{r} 
#| code-line-numbers: "2,6"

ui <- fluidPage(
  textOutput(outputId = "text")
)

server <- function(input, output, session) {
  output$text <- renderText({
  })
}
```

## Using Inputs

- This works the same with inputs.
- The server can access an object named `input` that will contain the value of all the input objects defined in the UI.
- This object is **list-like**, just like the output object. Each input value has its own entry in this list, e.g. `input$number`.

## 

Combining the input and output objects, we get a simple app that displays the square of a number. ðŸ‘

```{r}
#| eval: false
library(shiny)

ui <- fluidPage(
  numericInput(inputId = "number", 
               label = "Enter a number", 
               value = 0),
               
  textOutput(outputId = "text")
)

server <- function(input, output, session) {
  output$text <- renderText({
    input$number^2
  })
}

shinyApp(ui, server)
```

## Your Turn! 

- Create a Shiny app that displays the square of a number
- Try entering different numbers manually
- Try using the up and down arrows
- Adjust the server logic to display the cube of a number
- Try other calculations

```{r}
#| echo: false

countdown(minutes = 10, 
          id = "timer1", 
          bottom = "10%",
          right = "10%")
```

## Eager Output

An output is **eager**: it will update as soon as the input changes.

<br>

This **eagerness** is handy: you don't need to worry about updating the output when the input changes.

<br>

But what if you want to trigger the calculation only when you want?

## Adding a Reactivity Modifier

You could use an `actionButton` as an **event**:

```{r} 
#| code-line-numbers: "8-9,15-17"

library(shiny)

ui <- fluidPage(
  numericInput(inputId = "number", 
               label = "Enter a number", 
               value = 0),
               
  actionButton(inputId = "button", 
               label = "Click me"),
               
  textOutput(outputId = "text")
)

server <- function(input, output, session) {
  output$text <- renderText({
    input$number^2
  }) |> bindEvent(input$button)
}

shinyApp(ui, server)
```

## Try It Yourself! 

- Add an `actionButton` to your app
- Adjust the text on the button to read "Calculate"
- Make sure it functions as intended
- Go to the [getting started](https://shiny.posit.co/r/getstarted/shiny-basics/lesson3/) lesson on control widgets
- See if you can control the calculation with a checkbox instead of a button
- What is wrong with this approach?

## Acknowledgements

- Parts of this presentation were adapted from Veerle van Leemput's [Shiny 101: The Modular App Blueprint](https://github.com/hypebright/shinyconf2024-shiny101)
