---
title: "Coding Assignment 1"
---

## Overview

For this assignment, you are going to download some variables using the `wbstats` and `vdemlite` packages. Then you are going to wrangle these data and merge the two data sets into one and analyze how they relate to women's representation in national parliaments. Do your work for each step in the code chunks provided. 

Here is a setup code chunk. You can load all of your packages here or as you go along in the code chunks below using the `library()` function. However, note that you really only need to load a library once per document. 

```{r}

```

:::{.callout-warning}
Please refrain from *installing* any packages in your code chunks because this will also install them every time the code chunk is run or your document is rendered. This may not be desirable for you or whoever is looking at or running your code (e.g. me).
:::

## Step 1: Download data from V-Dem (20pts)

*Use the V-Dem [codebook](https://www.v-dem.net/data/reference-documents/) or the `searchdem()` function in `vdemlite` to identify one measures of democracy and one indicator related to women's empowerment. Now use `fetchdem()` to download the most recent 25 years or so of data for your analysis for all countries in the dataset. Make sure to load the packages that you need and `glimpse()` the data or `View()` it to make sure that it downloaded properly.* 

```{r}

```

## Step 2: Download data from the World Bank (20 pts)

*Next, download the variable on women's represenation that we used in Module 1.2 ("SG.GEN.PARL.ZS") and at least one additional measure related to women's empowerment. Go to the [WDI](https://databank.worldbank.org/source/world-development-indicators) site the wb_search() function to identify relevant variables. Download data for the same time period as you did for the V-Dem data for all countries.* 

::: {.callout-important}
Try to make sure you download indicators with enough data to conduct your analysis. You won't get great results if there are too many NAs. 
:::

```{r}

```

## Step 3: Merge the data (20 pts)

*Now add country codes using the `countrycode` package and merge the data using `left_join()`.* 

```{r}

```

## Step 4: Summarize your combined data set (20 pts)

*Use `group_by()`, `summarize()` and `arrange()` to glean insights about your data. For example, how do regions compare on mean values of women's representation and how do these values relate to the values of other variables in your data set? Which countries stand out in terms of women's representation? Etc.* 

```{r}

```

## Step 5: Interpret your data (20 pts)

*Write a paragraph or so relating your insights to the Norris reading about electoral reform and women's representation. Picking a handful of country cases from your data, how does your analysis relate to her arguments about the relevance of electoral systems, statutory quotas, reserved seats or voluntary quotas?* 

::: {.callout-warning}
Don't write your paragraphs or any interpretation or analysis in a code chunk. Use markdown text. Quarto is designed to be a literate programming tool, so you can write your analysis in markdown text and include code chunks when necessary to support your argument.
:::

